---
title: "Lab 3"
author: "Chloe Swick"
date: "2023-01-26"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(jtools)
library(AICcmodavg)
```

# Pseudocode 

* examine our data, tables, summary stats 
* identify a question 
* wrangle data if necessary 
* identify some candidate models 
* select among candidates models using AIC/BIC 
* select among candidate models using k-fold cross validation 
* select among candidate models using area under receiver operating characteristic curve 


```{r}
GGally::ggpairs(penguins %>% select(species, bill_length_mm:sex),
                aes(color = species))
## quick way to look at information about data
```

```{r}
class(penguins$species)
levels(penguins$species)
```

```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c('Adelie', 'Chinstrap')) %>% 
  mutate(species = fct_drop(species)) %>% 
  select(-year) %>% 
  drop_na()
levels(adelie_chinstrap$species)
```


== means a match, when using with a vector, it matches the pattern as well 
%in% means a match to either or any of the items in vector 


```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm))+
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~species)

ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~species) # facet_wrap makes two different graphs per how many categoeries of the vairable you choose, in this example, we have a plot of adelie, and a plot for chinstrap 
```

# for fun 
```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex, shape = species)) +
  facet_wrap(~island)
```



# binary logistic regression 
```{r}
f1 <- species ~ body_mass_g + flipper_length_mm + sex 

ad_chin_blr1 <- glm(formula = f1, data = adelie_chinstrap, 
                    family = 'binomial')
ad_chin_blr1
summary(ad_chin_blr1)

blr1_tidy <- tidy(ad_chin_blr1) # creates a table of summary statistics 
```

















